A Closer Look at Network Flow on Evolving Graphs

### 2018-05-15 paper.tex
include more experiment code

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n8) at (-10,5) {C};
  \node[circle,draw=red1, minimum size=0.2cm] (n10) at (-8,10) {A};
  \node[circle,fill=green1, minimum size=0.2cm] (n6) at (3,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {A};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {B};
  \node[circle,draw=red1, minimum size=0.2cm] (n12) at (6,5)  {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n14) at (8,10) {A};
  \foreach \from/\to in {n7/n8, n1/n4, n4/n1, n1/n6, n14/n11}
   \draw[every edge,->] (\from) -- (\to);
 %\draw[dotted,->](n4) -- (n12);
% \draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $2$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $100$};
    \end{tikzpicture}
\end{center}
\caption{An evolving directed graph with 3 time stamps $1$, $2$ and $100$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows.}
\label{fig:katz_eg3}
\end{figure}

@@

\begin{lstlisting}
function temporal_katz(g::AbstractEvolvingGraph,
  start::TimeNode; alpha = 0.2, max_level = 10)
    score = 0.
    v = start
    fronter = [v]
    level = 0
    while level < max_level
        next = []
        for u in fronter
            for v in forward_neighbors(g, u)
                push!(next, v)
                td = temporal_distance(start, u)
                d = td + level
                score += alpha^d
            end
        end
        fronter = next
        level += 1
    end
    return score
end
\end{lstlisting}

@@

\subsection{Temporal Communicability Betweenness Centrality}

% communicability from first principle

@@

\begin{wz}
  We may also consider betweenness centrality.
\end{wz}

Recall that betweenness centrality of a node in a static graph is defined by counting the fraction of shortest paths that traverse it.
We could extend
betweenness centrality on evolving graphs by replacing shortest paths with temporal shortest paths \cite{nicosia13}.
\begin{equation}
  \label{eq:bet1}
C_i^{betweenness} = \sum_{j \in V}\sum_{k \in V, k \ne j}\frac{\alpha_{jk}(i)}{\alpha_{jk}},
\end{equation}
where $\alpha_{jk}$ represents the number of temporal shortest paths from node $j$ to node $k$ and $\alpha_{jk}(i)$ represents the number of temporal shortest paths from node $j$ to node $k$ that pass through the node $i$.

One may want to take into account the length of time a node passes a message. The temporal betweenness centrality of the node $i$ at time stamp $t_m$ is defined as
\begin{equation}
  \label{eq:bet2}
  C_i^{betweenness}(t_m) = \sum_{j\ne i}\sum_{k\ne j, k\ne i}\frac{U_{j,k,t_m}(i)}{\alpha_{j,k}},
\end{equation}
where, as before, $\alpha_{j,k}$ represents the number of temporal shortest paths from $j$ to $k$ and $U_{j,k,t_m}(i)$ represents the number of temporal shortest paths from $j$ to $k$ in which node $i$ is traversed from
the walk in the time stamp $t_m$ or a previous time stamp $t_{m-1}$. Then
average temporal betweenness of node $i$ is then defined as
$$
  C_i^{betweenness} = \frac{1}{M}\sum_m C_i^{betweenness}(t_m).
$$
Note \eqref{eq:bet1} only takes account of static edges while \eqref{eq:bet2} takes account of both static edges and causal edges.

% closeness centrality
Similar to betweenness centrality, let $d_{ij}$ be the length of a temporal shortest walk from $i$ and $j$ in an evolving graph. Then
the closeness centrality of a node $i$ is defined
$$
C_i^{closeness} = \frac{N-1}{\sum_j d_{ij}}.
$$
Note that this depends on how we define the length of temporal shortest walk. The closeness centrality can takes account of only static edges, only causal edges, or both.

@@

We derive two generalization of PageRank on evolving graphs.
The first idea is based on reverse temporal flows.

@@

We consider two generalisation of pagerank: a block matrix one and a dense matrix one that is based on probability

We believe we shouldn't not just penolaize old walks but the time span between two edges.

@@

@@

@@

We construct the evolving coauthor network from Etymo (\url{https://etymo.io}).
We collected $1566$ publications published by Journal of Machine Learning Research (JMLR) from $2001$ to $2017$ with authors involved. We regard each year
as a time stamp and there are $17$ time stamps in total. At each time stamp, we
create a undirected edge (or two directed edges) between two authors if they have coauthored at least one paper.
(add results here.)

A second network is constructed by linking leading authors in different papers whose
ideas are similar. The resulting network represent the connectivity of ideas in all the research papers in the database.

(add results here.)

% \subsection{Paper Keyword Tag Network}
%
% We tag our research articles by subjects.
%
% We can construct a time-dependent tag network by link tags in each articles. For example,
% suppose a paper $p_1$ has three tags $t_1$, $t_2$, and $t_3$, where $t_1$ is a superclass of $t_2$ and $t_3$.
% We link $t_2$ and $t_3$ in both direction and create a link from $t_1$ to $t_2$ and a link from $t_1$ to $t_3$.
% The time stamp of these edges is the published date of the paper.

@@

To write later.

>>>>

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n8) at (-10,5) {C};
  \node[circle,draw=red1, minimum size=0.2cm] (n10) at (-8,10) {A};
  \node[circle,fill=green1, minimum size=0.2cm] (n6) at (3,7) {B};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {A};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {B};
  \node[circle,draw=red1, minimum size=0.2cm] (n12) at (6,5)  {C};
  \node[circle,fill=green1, minimum size=0.2cm] (n14) at (8,10) {A};
  \foreach \from/\to in {n7/n8, n1/n4, n4/n1, n1/n6, n14/n11}
   \draw[every edge,->] (\from) -- (\to);
 %\draw[dotted,->](n4) -- (n12);
% \draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $10$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $100$};
    \end{tikzpicture}
\end{center}
\caption{An evolving directed graph with 3 time stamps $1$, $10$ and $100$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows.}
\label{fig:katz_eg3}
\end{figure}

@@

\begin{lstlisting}[caption={Temporal Katz Centrality of Single Node},label={lst:katz}]
function temporal_katz(g::AbstractEvolvingGraph,
  start::TimeNode; alpha = 0.2, max_level = 10)
    score = 0.
    v = start
    fronter = [v]
    level = 0
    while level < max_level
        next = []
        for u in fronter
            for v in forward_neighbors(g, u)
                push!(next, v)
                td = temporal_distance(start, u)
                d = td + level
                score += alpha^d
            end
        end
        fronter = next
        level += 1
    end
    return score
end
\end{lstlisting}

@@

\subsubsection{Temporal Resolvent Betweenness Centrality}

Betweenness centrality measures the importance of a node as the ability to facilitate the communication between other nodes in the network. Commonly it is defined based on the shortest paths. However, key messages do not necessarily follow the shortest paths and so it makes senses to consider walks instead.
Recall the $(i,j)$ entry of the resolvent matrix $(I - \alpha A)^{-1}$ provide information about the communication from $i$ to $j$. We could therefore define the resolvent betweenness for node $v$ as
\begin{equation}
  \sum_{i}\sum_j \frac{(I-\alpha A)^{-1}_{i,j} - (I - \alpha (A - E_v))^{-1}_{i,j}}{(I - \alpha A)^{-1}_{i,j}}, \quad i \ne j, i \ne v, j \ne v.
\end{equation}
where $E_v$ has non-zeros only in row and column $v$, and in the $v$th row and column has $1$ where $A$ has $1$.
The matrix $A - E_i$ represents a graph with all edges involving the node $v$ removed.
For $(I-\alpha A)^{-1}_{i,j}$ we could simply modify Listing \ref{lst:katz} so that it only accumulate scores when the end node is $j$. For the $(I - \alpha (A-E_v))^{-1}_{i,j}$ part, we can apply the above the same algorithm on a graph with node $v$ and related edges removed.
\begin{lstlisting}
function temporal_resolvent_betweenness(g1::EvolvingGraph,
  g2::EvolvingGraph, v::TimeNode; alpha = 0.2, k = 10)
    r = 0.
    ns = active_nodes(g1)
    for i in ns
        for j in ns
            if i != j && i != v && j != v
                score1 = temporal_katz(g1, i, j, alpha = alpha, k = k)
                score2 = temporal_katz(g2, i, j, alpha = alpha, k = k)
                r += score1 == 0.0 ? 0 : (score1 - score2)/score1
            end
        end
    end
    return r
end
\end{lstlisting}
For example, by removing node $C$ a time stamp $2$ from Figure \ref{fig:katz_eg2} we find
the temporal resolvent betweenness score is $6.09$. While the temporal resolvent betweenness score
of node $A$ at time stamp $3$ is $5.0$. This shows $(C,2)$ is more important than $(A,3)$.

Now consider Figure \ref{fig:katz_eg3}. If we remove $(C,2)$, the temproal resolvent betweenness score is $4.20$.
The analysis for temporal communicability betweenness centrality is similar.

@@

% closeness centrality
Let $d_{ij}$ be the length of a temporal shortest walk from $i$ and $j$ in an evolving graph. Then
the closeness centrality of a node $i$ is defined
$$
C_i^{closeness} = \frac{N-1}{\sum_j d_{ij}}.
$$
Note that this depends on how we define the length of temporal shortest walk. The closeness centrality can takes account of only static edges, only causal edges, or both.
The definition of temporal closeness centrality is similar to temporal katz. The only difference is that here we traverse the shortest paths not just walks on evolving graphs.

\begin{lstlisting}
function temporal_closeness(g::AbstractEvolvingGraph, start::TimeNode)
  v = start
  level = Dict(v => 0)
  i = 1
  fronter = [v]
  while length(fronter) > 0
      next = []
      for u in fronter
          for v in forward_neighbors(g, u)
              if !(v in keys(level))
                  td = temporal_distance(start, u)
                  level[v] = i + td
                  push!(next, v)
              end
          end
      end
      fronter = next
      i += 1
  end

  total_scores = sum(values(level))
  return total_scores > 0. ? (length(level) - 1)/total_scores : 0.
end
\end{lstlisting}
For Figure \ref{fig:katz_eg2} the scores are
\begin{lstlisting}
Dict{Any,Any} with 7 entries:
TimeNode(B, 3) => 0.0
TimeNode(B, 1) => 0.27027
TimeNode(A, 3) => 1.0
TimeNode(B, 2) => 1.0
TimeNode(C, 1) => 0.259259
TimeNode(A, 2) => 0.538462
TimeNode(C, 2) => 0.416667
\end{lstlisting}
For Figure \ref{fig:katz_eg3}, the scores are
\begin{lstlisting}
Dict{Any,Any} with 7 entries:
TimeNode(B, 100) => 0.0
TimeNode(A, 2)   => 0.0636364
TimeNode(A, 100) => 1.0
TimeNode(C, 2)   => 0.0458716
TimeNode(B, 1)   => 0.0746269
TimeNode(C, 1)   => 0.0564516
TimeNode(B, 2)   => 1.0
\end{lstlisting}

The analysis for temporal communicability betweenness centrality is similar.

@@

We derive a generalization of PageRank on evolving graphs.
The first idea is based on reverse temporal flows and block adjacency matrix.

@@

We derive the block google matrix as a combination of block matrix $A$ and its transpose $A^T$.

The block google matrix is defined as

\begin{lstlisting}
function block_google_matrix(g::IntAdjacencyList; alpha = 0.85, balance = 0.75)
  A = full(block_adjacency_matrix(g))
  A = balance * A + (1-balance) * A'
  N, N = size(A)
  p = ones(N)/N
  dangling_weights = p
  dangling_nodes = find(x->x==0, sum(A,2))

  for node in dangling_nodes
      A[node,:] = dangling_weights
  end
  A ./= sum(A,2)
  return alpha * A .+ (1- alpha) * p
end
\end{lstlisting}

We find the ranking for Figure \ref{}

\begin{lstlisting}
9-element Array{Complex{Float64},1}:
 0.12278
0.254264
0.0533627
 0.36275
0.350713
0.406792
0.0533627
0.524112
0.468848
\end{lstlisting}

@@

We test the influence of time on evolvign graph centralities.

@@

Note the random evolving graphs are conducted using \texttt{IntAdjacencyList}, which gives better performance for
forward\_neighbors.

We generate a random evolving graph with $100$ nodes and $9771$ static edges and $5$ time stamps.
Here are the top $10$ nodes we find using the Katz Centrality

Since we only are interested in the top nodes.

\begin{tabular}{ l | c }
  \hline
  node & score \\
  334 & 0.152288 \\
  9 & 0.137403 \\
  428 & 0.126793 \\
  164 & 0.121582 \\
  250 & 0.117862 \\
  35 & 0.117627 \\
  291 & 0.105793 \\
  414 & 0.0975953 \\
  123 & 0.096394 \\
  478 & 0.0945754 \\
  \hline
\end{tabular}

We find identify the top $10$

We test temporal Katz centrality,  for top $20$
temporal closeness, and temporal pagerank.
by varing the influence of temporal distance.

Katz score | temproal katz at with time influence 0.1 | temproal katz with time influence 1.0

Closeness with time influence 0.1 | closeness with time influence 1.0

PageRank with time influence 0.1 | PageRank with time influence 1.0

@@

We construct the evolving coauthor network from Etymo (\url{https://etymo.io}).
We collected $1566$ publications published by Journal of Machine Learning Research (JMLR) from $2001$ to $2017$ with authors involved. We regard each year
as a time stamp and there are $17$ time stamps in total. At each time stamp, we
create a undirected edge (or two directed edges) between two authors if they have coauthored at least one paper.

Some experiments as above.

Katz score | temproal katz at with time influence 0.1 | temproal katz with time influence 1.0

Closeness with time influence 0.1 | closeness with time influence 1.0

PageRank with time influence 0.1 | PageRank with time influence 1.0

@@

We describe in details the generalizartion of centrality algorithms for evolving graphs.
We introduce the generalization of PageRank centrality.
